Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-02-10 Manuscript/Tmatrix.tex
moving figures
data for molecules
savig work

\DeclareSIUnit\angstrom{\text{\AA}}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} which is freely available at \url{https://github.com/pfloos/QuAcK}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here.
Although the dynamical correction is computed in the dTDA throughout, the zeroth-order excitonic Hamiltonian [see Eq.~\eqref{eq:BSE}] is always the ``full'' BSE static Hamiltonian, i.e., without TDA.

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of the hydrogen molecule}
\label{sec:h2}
%%%%%%%%%%%%%%%%%%%%%%
As a first didactical example, we consider the lowest singlet and triplet excited states of the hydrogen molecule \ce{H2} and the variation of their respective vertical transition energies upon dissociation.
The excitation energies associated with these low-lying excited states are represented in Fig.~\ref{fig:H2} as a function of the internuclear distance $R_{\ce{H-H}}$ at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
The variation of the HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:H2_gap}
This shows that, as already observed in the Hubbard dimer \cite{Romaniello_2012} and in molecular systems, \cite{Zhang_2017} the $G_0W_0$ and $G_0T_0$ quasiparticle energies are similar near the Fermi level.

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of beryllium hydride}
\label{sec:beh2}
%%%%%%%%%%%%%%%%%%%%%%
As a second example, we consider the symmetric dissociation of the linear molecule beryllium hydride (\ce{BeH2}), a system for which one can assume that the screening may play a more important role than in the previous example.
The variation of the lowest singlet and triplet excitation energies as a function of the distance $R_{\ce{Be-H}}$ is shown in Fig.~\ref{fig:BeH2}, while the quasiparticle energies of the frontier orbitals and the associated (fundamental) gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:BeH2_gap}.
Again, one notes that the $G_0W_0$ and $G_0T_0$ quasiparticle energies are very similar near the Fermi level.
Therefore, one can safely assume that any significant variation of the excitation energies computed within the $GW$- and $GT$-based formalisms originates mainly from their distinct kernel.
The excitation energies computed with the dynamical schemes, dBSE@$G_0W_0$ and dBSE@$G_0T_0$, are reported as thin solid lines.
Here, one can show that dynamical corrections improves in most cases the agreement between BSE and FCI.

@@

@@

Table \ref{tab:mol} reports singlet and triplet vertical excitation energies for various molecular systems at equilibrium geometry.
In addition to the BSE-based models studied in the present manuscript, we have selected well-known methods \titou{(CIS, CIS(D), TDHF and FCI)}.

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the cc-pVDZ basis set.
	The corresponding experimental equilibrium geometries have been extracted from the computational chemistry comparison and benchmark database (CCCBDB) at \url{https://cccbdb.nist.gov}.}
	\label{tab:mol}
	\begin{ruledtabular}
		\begin{tabular}{lcccccccc}
					&			&	\mc{7}{c}{Methods}																\\
			\cline{3-9}
			Mol.	&	State	&	CIS		&	CIS(D)	&	TDHF	&	BSE@$GW$	&	dBSE@$GW$	&	BSE@$GT$	&	dBSE@$GT$	\\
			\hline
			\ce{H2}	&			&	\\
			\ce{LiH}	&			&	\\
			\ce{LiF}	&			&	\\
			\ce{BF}	&			&	\\
			\ce{CO}	&			&	\\
			\ce{N2}	&			&	\\
			\ce{F2}	&			&	\\
			\ce{FH}	&			&	\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%

>>>>

\newcommand{\pps}{\pi \to \pi^*}
\newcommand{\nps}{n \to \pi^*}
\DeclareSIUnit\angstrom{\text{\AA}}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Computational details}
\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%
The present formalism has been implemented in the electronic structure package \texttt{QuAcK} \cite{QuAcK} which is freely available at \url{https://github.com/pfloos/QuAcK}. 
We consider here only systems with closed-shell singlet ground states. 
Thus, the $GW$ and $GT$ calculations are performed by considering a (restricted) HF starting point and standard gaussian basis sets (defined with cartesian functions) are employed.
Note that all quasiparticle energies which are obtained via Eq.~\eqref{eq:G0T0} are corrected in the same way.
Finally, the infinitesimal $\eta$ is set to zero for all calculations.
The ev$GT$ and qs$GT$ schemes have been also implemented but are not considered here.
Although the dynamical correction is computed in the dTDA throughout, the zeroth-order excitonic Hamiltonian [see Eq.~\eqref{eq:BSE}] is always the ``full'' BSE static Hamiltonian, i.e., without TDA. 
Reference full configuration interaction (FCI) calculations have been performed with {\QP}. \cite{Garniron_2019}

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of the hydrogen molecule}
\label{sec:h2}
%%%%%%%%%%%%%%%%%%%%%%

As a first didactical example, we consider the lowest singlet and triplet excited states of the hydrogen molecule \ce{H2} and the variation of their respective vertical transition energies upon dissociation.
The excitation energies associated with these low-lying excited states are represented in Fig.~\ref{fig:H2} as a function of the internuclear distance $R_{\ce{H-H}}$ at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
The variation of the HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:H2_gap}
This shows that, as already observed in the Hubbard dimer \cite{Romaniello_2012} and in molecular systems, \cite{Zhang_2017} the $G_0W_0$ and $G_0T_0$ quasiparticle energies are similar near the Fermi level.

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of beryllium hydride}
\label{sec:beh2}
%%%%%%%%%%%%%%%%%%%%%%

@@

As a second example, we consider the symmetric dissociation of the linear molecule beryllium hydride (\ce{BeH2}), a system for which one can assume that the screening may play a more important role than in the previous example.
The variation of the lowest singlet and triplet excitation energies as a function of the distance $R_{\ce{Be-H}}$ is shown in Fig.~\ref{fig:BeH2}, while the quasiparticle energies of the frontier orbitals and the associated (fundamental) gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:BeH2_gap}.
Again, one notes that the $G_0W_0$ and $G_0T_0$ quasiparticle energies are very similar near the Fermi level.
Therefore, one can safely assume that any significant variation of the excitation energies computed within the $GW$- and $GT$-based formalisms originates mainly from their distinct kernel.
The excitation energies computed with the dynamical schemes, dBSE@$G_0W_0$ and dBSE@$G_0T_0$, are reported as thin solid lines.
Here, one can show that dynamical corrections improves in most cases the agreement between BSE and FCI.
All calculations are performed with the cc-pVDZ basis.

@@

Table \ref{tab:mol} reports singlet and triplet vertical excitation energies for various molecular systems at equilibrium geometry.
In addition to the BSE-based models studied in the present manuscript, we have selected well-known methods, namely CIS, CIS(D), TDHF, and FCI (taken as reference).

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the aug-cc-pVDZ basis set.
	The corresponding equilibrium geometries as well as the reference FCI values have been extracted from the QUEST database. \cite{Veril_2021}}
	\label{tab:mol}
	\begin{ruledtabular}
		\begin{tabular}{llccccccccc}
					&			&		&	\mc{8}{c}{Methods}																\\
			\cline{4-11}
			Mol.	&	State	&	Nature						&	CIS		&	CIS(D)	&	TDHF	&	BSE@$G_0W_0$	&	dBSE@$G_0W_0$	&	BSE@$G_0T_0$	&	dBSE@$G_0T_0$	&	FCI	\\
			\hline
			\ce{H2O}	&	$^1B_1(n \to 3s)$		&	Ryd.	&	8.66	&	6.61	&	8.61	&	7.70	&	7.62	&	6.84	&	6.80	&	7.53\\
						&	$^1A_2(n \to 3p)$		&	Ryd.	&	10.34	&	8.33	&	10.30	&	9.41	&	9.34	&	8.60	&	8.56	&	9.32\\
						&	$^3B_1(n \to 3s)$		&	Ryd.	&	7.98	&	6.69	&	7.86	&	7.23	&	7.09	&	6.30	&	6.23	&	7.14\\
						&	$^3A_2(n \to 3p)$		&	Ryd.	&	10.01	&	8.70	&	9.89	&	9.22	&	9.11	&	8.36	&	8.29	&	9.14\\
			\ce{HCl}	&	$^1\Pi$					&	CT		&	8.36	&	7.87	&	8.31	&	8.04	&	7.95	&	7.42	&	7.38	&	7.82\\
			\ce{HC#CH}	&	$^1\Sigma_u^-(\pps)$	&	Val.	&	6.28	&	6.29	&	5.91	&	7.27	&	6.96	&	5.64	&	5.55	&	7.20\\
						&	$^1\Delta_u(\pps)$		&	Val.	&	6.62	&	6.97	&	6.44	&	7.61	&	7.33	&	5.87	&	5.81	&	7.51\\
						&	$^3\Sigma_u^+(\pps)$	&	Val.	&	4.52	&	5.75	&	2.19	&	5.63	&	5.12	&	4.02	&	3.83	&	5.50\\
						&	$^3\Delta_u(\pps)$		&	Val.	&	5.45	&	6.65	&	4.52	&	6.52	&	6.13	&	4.87	&	4.74	&	6.46\\
			\ce{H2C=NH}	&	$^1A''(\nps)$			&	Val.	&	5.86	&	4.46	&	5.69	&	6.16	&	5.85	&	4.12	&	4.04	&	5.25\\
						&	$^3A''(\nps)$			&	Val.	&	4.71	&	4.72	&	4.39	&	5.49	&	4.96	&	3.25	&	3.11	&	4.63\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%
